Flaring star-planet interactions in Kepler and TESS

### 2023-06-03 src/tex/ms.tex
[showyourwork] overleaf sync

If we picture the magnetic field lines connecting the planet to the star in the sub-Alfv\'enic zone, we expect that the stellar footpoint of these lines will move along with the planet. Flares triggered by the planet will occur at a preferred orbital phase, namely whenever the footpoint is visible to the observer. Depending on the system's viewing geometry, i.e., the inclination of the rotation axis, the spin-orbit (mis-)alignment, and the extent and configuration of the large scale magnetic field -- the footpoint could be always in view, or never. In expectation, however, in star-planet systems with favorable viewing geometries and magnetic field strengths and configurations, the footpoint visibility will be modulated in phase with the planetary orbit~(Fig.~\ref{}). The resulting phase preference of planet-induced flares introduces a deviation from a random distribution of flares. Space missions like Kepler and TESS provide light curves that cover all phases of the planet's orbit, so that phase-correlated planet-induced flares can be detected against a background of uncorrelated intrinsic flaring.

\begin{figure*}[ht!]
\script{paper_illustrative_sketch_of_system_geometries.py}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/system_geometries.png}
        \caption{
         Simplified magnetic star-planet interaction geometries.\textbf{ Panel a: Fully aligned system along the line of sight. }Magnetic, rotational, and orbital axis are aligned. The interaction footpoint moves along the black line, and is visible on the red line, i.e. during $50\%$ of the planet's orbit. Flares induced along this line will result in excess flares when the planet is in front of the star. \textbf{Panel b: Orbital misalignment.} The rotational and magnetic axis are aligned, but the planet is highly misaligned. The footpoint of interaction moves between the two horizontal black lines, and is still visible $50\%$ of the time, resulting in modulated flaring with orbital phase. However, the footpoint passes different latitudes on the star, which correspond to different dipole field strengths, which may influence the power of interaction. \textbf{Panel c: Inclined fully aligned system.} The red interaction footpoint line is visible $100\%$ of the time. No modulation due to SPI can observed, even if flares are triggered by the planet. The sketch only shows the interaction footpoint region on the Northern hemisphere of the star, but they may appear on both.
        }
        \label{fig:sketch}
    \end{centering}
\end{figure*}

>>>>

If we picture the magnetic field lines connecting the planet to the star in the sub-Alfv\'enic zone, we expect that the stellar footpoint of these lines will move along with the planet. Flares triggered by the planet will occur at a preferred orbital phase, namely whenever the footpoint is visible to the observer. Depending on the system's viewing geometry, i.e., the inclination of the rotation axis, the spin-orbit (mis-)alignment, and the extent and configuration of the large scale magnetic field -- the footpoint could be always in view, or never. In expectation, however, in star-planet systems with favorable viewing geometries and magnetic field strengths and configurations, the footpoint visibility will be modulated in phase with the planetary orbit~(Fig.~\ref{fig:sketch}). The resulting phase preference of planet-induced flares introduces a deviation from a random distribution of flares. Space missions like Kepler and TESS provide light curves that cover all phases of the planet's orbit, so that phase-correlated planet-induced flares can be detected against a background of uncorrelated intrinsic flaring.

\begin{figure*}[ht!]
\script{paper_illustrative_sketch_of_system_geometries.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/system_geometries.png}
        \caption{
         Simplified magnetic star-planet interaction geometries.\textbf{ Panel a: Fully aligned system along the line of sight. }Magnetic, rotational, and orbital axis are aligned. The interaction footpoint moves along the black line, and is visible on the red line, i.e. during $50\%$ of the planet's orbit. Flares induced along this line will result in excess flares when the planet is in front of the star. \textbf{Panel b: Orbital misalignment.} The rotational and magnetic axis are aligned, but the planet is highly misaligned. The footpoint of interaction moves between the two horizontal black lines, and is still visible $50\%$ of the time, resulting in modulated flaring with orbital phase. However, the footpoint passes different latitudes on the star, which correspond to different dipole field strengths, which may influence the power of interaction. \textbf{Panel c: Inclined fully aligned system.} The red interaction footpoint line is visible $100\%$ of the time. No modulation due to SPI can be observed, even if flares are triggered by the planet. The sketch only shows the interaction footpoint region on the Northern hemisphere of the star, but they may appear on both.
        }
        \label{fig:sketch}
    \end{centering}
\end{figure*}
